---
title: "Mini-Project 2"
author: "Yifan Ma & Meghan Suslovic"
date: "March 7, 2018"
output: 
  html_document:
    code_folding: hide

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE, echo=TRUE)
```


```{r, include=FALSE}
load("house_elections.rda")
load("candidates.rda")
load("committees.rda")
load("contributions.rda")
```

```{r}
library(tidyverse)
library(maps)
library(mapdata)
library(ggthemes)
library(RColorBrewer)
```


**Money and Power: Swing States**

Given the upcoming special elections this fall, a lot of attention is being paid to special elections, like the most recent house election in Pennsylvania in which a Democrat narrowly beat a Republican in a district that Trump won in 2016. More often, this attention is directed towards swing states (i.e. states that can swing Democratic or Republican in any given race). We were curious to see if campaign contributions were higher in these toss up states given the high stakes associated with winning elections in these states. We found that average contribution amounts were much higher for campaigns that ultimately won their general elections in 2012 and campaigns in swing states also had a lot more money pouring into them.    

To start, we joined the contributions data frame to the house election data, so the transaction amounts is directly linked to the information of  winners and losers in all the house elections that occurred in 2012. The total amount from all winners and losers are summed up and mean is taken. The final plot took form in a bar plot for directness. 

Figure 1 shows that the winning 2012 house election campaigns received an average of $1,040,503 while losing campaigns received only $347,372, suggesting the average $693,131 extra the winning campaigns received from donors helped push them towards victory. More broadly, this signals to donors that their contributions will impact the outcome of the election, creating the incentive to pour ever increasing amounts of money into campaigns. Further research should examine whether this trend holds true for elections at the local and state level and in other federal elections.

```{r}
committees2<- committees %>%
  select(cmte_id:connected_org_name)

all <- house_elections %>%
  left_join(contributions, by = c("fec_id" = "cand_id")) %>%
  group_by(fec_id,ge_winner) %>%
  summarise(S = sum(transaction_amt))%>%
  filter(!is.na(S))%>%
  group_by(ge_winner)%>%
  summarise(m = mean(S))

all <- all %>%
  mutate(nname = case_when( ge_winner == "N" ~ "Lost", ge_winner == "W" ~ "Won"))

ggplot(all, aes(x = nname, y = m))+
  geom_bar(stat = "identity")+
  ggtitle(label = "Figure 1: Mean Transaction Amount Received by Winning and Losing Campaigns", subtitle = "Source: FEC")+
  labs(x = "",  y = "Transaction Amount ($)")
  
```

From the joined data frame from house election and contribution, we grouped by states and summed the total transaction amount happened within that state in 2012. The sum of each state is then divided by the number of districts within the state. Then we use a barplot to present how much money transaction within each state. 

```{r}
d <- house_elections %>%
  group_by(state) %>%
  summarise(n = n_distinct(district))


state <- house_elections %>%
  left_join(contributions, by = c("fec_id" = "cand_id"))%>%
  group_by(state.x) %>%
  filter(!is.na(transaction_amt)) %>%
  summarise(S2 = sum(transaction_amt)) %>%
  left_join(d, by = c("state.x" = "state"))%>%
  mutate(mean_d = S2/n) %>%
  arrange(desc(mean_d))

ggplot(state, aes(x = state.x, y = mean_d))+
  geom_bar(stat = "identity")+
  ggtitle(label = "Figure 2: Mean Transaction Amounts Received by House Election Campaigns by State", subtitle = "Source: FEC")+
  scale_y_continuous(labels= scales::dollar)+
  labs(x = "State",  y = "Transaction Amount ($)")+
  theme(axis.text.x= element_text(angle=45)) 
  
  
```


```{r}
library(readxl)
state_a <- read_excel("state_a.xls")
states <- map_data("state")
states <- states %>%
  left_join(state_a, by = c("region" = "State")) %>%
  left_join(state, by = c("Abreviation" = "state.x"))


```

The final map is created by joining the mean transaction amount within each state to the every single state coordinates. Scale_fill_gradient is used to show the variation of the amounts. 

```{r}
base <- ggplot(data = states) + 
  geom_polygon(aes(x = long, y = lat, group = group, fill = mean_d), color = "white")+
  scale_fill_gradient(low = "light green", high = "dark green", name = "Mean transaction amount")+
  ggtitle(label = "Figure 3: Map of Mean Transaction Amounts Received by House Election Campaigns by State", subtitle = "Source: FEC")+
  theme(legend.position="right")+
  theme_void()+
  coord_fixed(1.3)

base
```


[Figure 4: 2012 Electoral College Map](http://www.rasmussenreports.com/var/plain/storage/images/media/commentary_images/sabato/sabato_map_1_may_7_2015/1104404-1-eng-US/sabato_map_1_may_7_2015.png) 
Source: Rasmussen Reports

From Figure 2, it is clear that New Hampshire (NH), Iowa (IA), and Nevada (NV) were the top three states that received the most house election campaign funding. For a more visual representation of the data, refer to Figure 3. The states with the darkest green shading, correspond to the states with the top funding amounts. There is considerable overlap with Figure 3 and Figure 4, an electoral map from 2012 that depicts the toss-up states in tan. New Hampshire, Iowa, and Nevada were all considered swing states in the 2012 presidential election, so it makes sense that donors were contributing large sums of money to elections in these states (particularly given that Figure 1 shows that winning campaigns received more money). Interestingly, these three states are also generally the first to host their presidential primaries/caucuses, meaning they receive even more attention from candidates and the media as an indicator for how the rest of the country will vote. Illinois is an outlier in this case because it was viewed as a dependly blue state, so it is unclear why so much money went to their house elections in 2012. It will be interesting to see if the patterns of winners and swing states receiving the most money will hold true during the upcoming midterm elections. If so, it seems there will be no incentive for big PACs to reign in their campaign donations if they want to continue shaping election outcomes.

**Footnotes:**
GitHub- https://github.com/yifan6210/sds192-mp2



